# Seattle Population Growth 2015 - 2020
```{r}

pop <- c(688156, 710510, 729601, 744955, 753675,776555)
year <- c(2015,2016,2017,2018,2019,2020)

a <- ggplot(data = NULL, aes(x = year, y = pop)) +
  geom_col(fill = '#8abbd0', width = 0.6, color = "black") +
  geom_smooth(color = "dodgerblue3", size = 2) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2015:2020), labels = c("2015", "2016", "2017", "2018", "2019","2020")) +
  geom_text(aes(label = format(pop, big.mark = ",")), position = position_stack(vjust = 1.05), size = 3) + # display the values count on the bar chart
  labs(x = NULL, y = "Population", title = "Seattle Population Growth 2015 - 2020") +
  theme_economist()
 ggplotly(a)

```

# Seattle Crime Rate 2015 - 2020
```{r}

pop <- c(688156, 710510, 729601, 744955, 753675,776555)
total_crime <- c(67882, 70437, 73100, 75325, 70289, 75790)
crime_rate <- round(total_crime/pop * 100000, digits = 0)

b <- ggplot(data = NULL, aes(x = year, y = crime_rate)) +
  geom_col(fill = '#8abbd0', width = 0.6, color = "black") +
  geom_smooth(color = "dodgerblue3", size = 2) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(2015:2020), labels = c("2015", "2016", "2017", "2018", "2019","2020")) +
  geom_text(aes(label = format(crime_rate, big.mark = ",")), position = position_stack(vjust = 1.05), size = 3) + # display the values count on the bar chart
  labs(x = NULL, y = "Reported Crimes per 100,000 residents", title = "Seattle Crime Rate 2015 - 2020") +
  theme_economist()
 ggplotly(b)
 
```

# Total Reported Crimes in Seattle 2015 - 2020
```{r, message = FALSE, warning =FALSE}

df1 <-  df %>% 
  #na.omit() %>% 
  group_by(year) %>% 
  summarise(count = n()) 
  
  g <- ggplot(df1,aes(x = year, y = count)) +
  geom_col(fill = '#8abbd0', width = 0.6, color = "black") +
  geom_smooth(color = "dodgerblue3", size = 2) +
  scale_x_continuous(breaks = c(2015:2020), labels = c("2015", "2016", "2017", "2018", "2019","2020")) +
  scale_y_continuous(labels =  scales::comma) +
  geom_text(aes(label = format(count, big.mark = ",")), position = position_stack(vjust = 1.05), size = 3) +
  labs(x = NULL, y = "Count", title = "Total Reported Crimes in Seattle 2015 - 2020") +
  theme_economist()
 ggplotly(g)  
 
```


# Categories of Crimes in Seattle 2015 - 2020

```{r, out.width='100%', fig.align="center"}

df %>% 
  filter(crime_category %in% c("PERSON", "PROPERTY", "SOCIETY")) %>% 
  group_by(crime_category, year) %>% 
  summarise(count = n()) %>% 
  
  ggplot(aes(x = year, y = count, fill = factor(crime_category, levels = c("SOCIETY","PERSON", "PROPERTY" )))) +
  guides(fill = guide_legend(title = "Crime Category")) +
  scale_x_continuous(breaks = c(2015:2020), labels = c("2015", "2016", "2017", "2018", "2019","2020")) +
  scale_y_continuous(labels = scales::percent) +
  geom_col(position = "fill", col = "gray30") +
  labs(x = NULL,y = NULL) + 
  scale_fill_brewer(palette = "Dark2") +
  scale_fill_manual(values=c("#A6D854", "#FC8D62", "#56B4E9"), labels = c("Society", "Person", "Property")) +
  theme_economist()
  
```


# Categories of Crimes in Seattle 2015 - 2020 (facet_wrap)
```{r, out.width='100%', fig.align="center"}

df %>% 
  filter(crime_category %in% c("PERSON", "PROPERTY", "SOCIETY")) %>% 
  group_by(crime_category, year) %>% 
  summarise(count = n()) %>% 
  
  ggplot(aes(x = year, y = count, fill = factor(crime_category, levels = c("PROPERTY","PERSON", "SOCIETY" )))) +
  guides(fill = guide_legend(title = "Crime Category")) +
  scale_x_continuous(breaks = c(2015:2020), labels = c("2015", "2016", "2017", "2018", "2019","2020")) +
  scale_y_continuous(labels = scales::comma) +
  geom_col(position = position_dodge2(), col = "gray30") +
  labs(x = NULL,y = "Count") + 
  scale_fill_brewer(palette = "Dark2") +
  scale_fill_manual(values=c("#56B4E9","#FC8D62","#A6D854"), labels = c("Property", "Person", "Society")) +
  theme_economist()
  
```


# Top 10 Most Dangerous Neighborhoods 2015 - 2020
```{r}

df %>% 
  count(area) %>% 
  mutate(area = fct_reorder(area, n)) %>%
  arrange(desc(n)) %>% 
  head(10) %>%   
  ggplot(aes(x = area, y = n)) +
  geom_col(fill = '#8abbd0', color = "black") +
  coord_flip() +
  scale_y_continuous(labels = scales :: comma) +
  geom_text(aes(label = format(n, big.mark = ",")), position = position_stack(vjust = 0.5), size = 3) + # display the values count on the bar chart
  labs(x = NULL, y = "Count of incidents") +
  theme_economist()
   
```
         
         
# Top 10 Crime Types in Seattle 2015 - 2020
```{r}

df %>% 
   count(parent_group) %>% 
  mutate(parent_group = fct_reorder(parent_group, n)) %>%
  arrange(desc(n)) %>% 
  head(10) %>% 

  ggplot(aes(x = parent_group, y = n)) +
  geom_col(fill = '#8abbd0') +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) + 
  geom_text(aes(label = format(n, big.mark = ",")), position = position_stack(vjust = 0.5), size = 3) + 
  labs(x = NULL, y = "Count", title = "Top 10 Crime Types in Seattle 2015 - 2020") +
  theme_economist()
  
```


# Top 10 Offense Types in Seattle 2015 - 2020 

```{r, out.width='100%', fig.align="center"}

df %>% 
  group_by(crime_category, offense) %>% 
  summarise(count = n()) %>% 
  mutate(offense = fct_reorder(offense, count)) %>% 
  arrange(desc(count)) %>% 
  head(10) %>% 

  
  ggplot(aes(x = reorder(offense, count), y = count, fill = factor(crime_category, levels = c("PROPERTY","PERSON", "SOCIETY" )))) +
  geom_col(lwd = 0.01, col = "black") +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) +
  geom_text(aes(label = format(count, big.mark = ",")), position = position_stack(vjust = 0.5), size = 3) + # display the values count on the bar chart
  guides(fill = guide_legend(title = "Crime Category")) +
  labs(x = NULL, y = "Count") +
  scale_fill_brewer(palette = "Dark2") +
  scale_fill_manual( values=c( "#56B4E9","#FC8D62", "#A6D854"), labels = c("Property", "Person", "Society")) +
  theme_economist()
  
```


# Drug Crimes in Seattle 2015 - 2020
```{r}

df_small9 <- df %>%
  filter(parent_group == "DRUG/NARCOTIC OFFENSES") %>%
  group_by(year) %>% 
  summarise(count = n())

 g <- ggplot(df_small9,aes(x = year, y = count)) +
  geom_col(fill = '#8abbd0', width = 0.6, color = "black") +
  geom_smooth(color = "dodgerblue3", size = 2) +
  scale_x_continuous(breaks = c(2015:2020), labels = c("2015", "2016", "2017", "2018", "2019","2020")) +
  scale_y_continuous(labels =  scales::comma) +
  geom_text(aes(label = format(count, big.mark = ",")), position = position_stack(vjust = 1.05), size = 3) + # display the values count on the bar chart
  labs(x = NULL, y = "Count", title = "Drug Crimes in Seattle 2015 - 2020") +
 theme_economist()
 ggplotly(g)  
 
```


# Top 5 Safest Neighborhoods in Seattle 2015 - 2020
```{r}

df %>% 
  count(area) %>% 
  mutate(area = fct_reorder(area, n)) %>%
  arrange(desc(n)) %>% 
  tail(10) %>% 
  ggplot(aes(x = area, y = n)) +
  geom_col(fill = '#8abbd0') +
  coord_flip() +
  scale_y_continuous(labels = scales ::comma) +
  geom_text(aes(label = format(n, big.mark = ",")), position = position_stack(vjust = 0.5), size = 3) + # display the values count on the bar chart
  labs(x = NULL, y = "Count of Incidents", title = "Top 5 Safest Neighborhoods 2015 - 2020") +
  theme_economist()
   
```


# Crimes by Day of Week in Seattle 2015 - 2020
```{r}

df %>% 
  ggplot(aes(x = day)) +
  geom_bar(fill = '#8abbd0', width = 0.7) +
  labs(x = NULL, y = "Count", title = "Crimes by Day of Week in Seattle 2015 - 2020") +
  scale_y_continuous(labels = scales::comma) +
  geom_bar(data = filter(df, day == "Fri"), fill = "dodgerblue3", width = 0.7) +
  theme_economist()
  
```


# Offense Types by Hour of Day
```{r}

df %>% 
  select(offense, year, month, day, hour) %>% 
  filter(offense %in% c("Theft From Motor Vehicle",
                        "Burglary/Breaking & Entering",
                        "Destruction/Damage/Vandalism of Property",
                        "All Other Larceny",
                        "Simple Assault",
                        "Shoplifting",
                        "Motor Vehicle Theft",
                        "Trespass of Real Property",
                        "Aggravated Assault",
                        "Intimidation")
         )
         

crime_by_hour <- function(data, offense_filter) 
{
  df_small1 <- data %>% 
    filter(offense == offense_filter) %>% 
    group_by(year,hour, crime_category) %>% 
    summarise(count = n()) 
  
  g <-ggplot(df_small1, aes(x = hour, y = count)) +
      geom_line(color = "dodgerblue3", size = 1.5) +
      facet_wrap(~year) +
      scale_x_continuous(breaks = seq(0,23, by = 1)) +
      labs( x= 'Hour' , y = "Count" ,title = paste(offense_filter, "by Hour of Day", sep = " ")) +
      theme_economist()

  return(g)
  }
  
offense_list <- c("Theft From Motor Vehicle",
                  "Burglary/Breaking & Entering",
                  "Destruction/Damage/Vandalism of Property",
                  "All Other Larceny",
                  "Simple Assault",
                  "Shoplifting",
                  "Motor Vehicle Theft",
                  "Trespass of Real Property",
                  "Aggravated Assault",
                  "Intimidation"
                  )

for (i in offense_list)
  print(crime_by_hour(df,i))
  
```


# Assault by Hour of Day 2015 - 2020
```{r}

df_small6 <- df %>%
  filter(offense %in% c("Simple Assault", "Aggravated Assault")) %>%
  group_by(year,hour) %>% 
  summarise(count = n())

  assault <- ggplot(df_small6, aes(x = hour, y = count, group = year)) +
  geom_line(color = "grey", size = 1)  +
  geom_line(data = filter(df_small6, year == 2020), color = "dodgerblue3", size = 1.3) +
  annotate(geom = "text", x = 23.5, y = 350, label = "2020", color = "dodgerblue3") +
  scale_x_continuous(breaks = seq(0,23, by = 1)) +
  labs( x= 'Hour' , y = "Count" ,title = "Assault by Hour of Day 2015 - 2020") +
  theme_economist() 
ggplotly(assault)

```


# Create interactive spatial maps
# Assault 
```{r}
my_title <- tags$b("Assault")
# ASSAULT OFFENSES
df_assault_address <- df %>% 
  select(parent_group, area, address, long, lat) %>% 
  filter(parent_group == "ASSAULT OFFENSES" ) %>% 
  group_by(area, address) %>% 
  summarise(count = n(), long = mean(long), lat = mean(lat)) %>%
  arrange(desc(count)) %>% 
  na.omit() %>% 
  head(50) 

icons <- makeIcon(iconUrl = 'https://cdn0.iconfinder.com/data/icons/topics/256/icon_crime-512.png',
                  iconWidth = 40, iconHeight = 40,
                  iconAnchorX = 20, iconAnchorY = 40)

df_assault_address %>%
  select(lat, long) %>% 
  leaflet() %>% 
  addTiles() %>% 
  addProviderTiles(provider = "Esri.WorldStreetMap",group = "World StreetMap") %>%
  addLayersControl(
    baseGroups = c("OSM (default)","World StreetMap"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addMarkers(icon =icons, popup = df_assault_address$address) %>% 
  addControl(my_title, position = "topright" )
```
